---
title: "From Data to Deliverable"
subtitle: "Day 1"  
author: "Hunter Glanz"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
# library(demoR)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
solarized_light()
```

class:center,middle,invert
# Prettifying your plots
---
# Tips

* Google is your friend!  Explore what other people have done.
* [https://www.r-graph-gallery.com/](https://www.r-graph-gallery.com/)
* [http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html)
* Multiple layers to the plot
* Look for the `+theme_` options
* Don't forget to think about colorblindness
---
```{r, fig.height = 5}
ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point()
```
---
```{r, fig.height = 5, warning = FALSE}
library(ggforce)

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point() +
  stat_ellipse() +
  ggtitle("Yay! Irises!") + 
  theme_light()
```

---
class:center,middle,invert
# Polishing and sharing your output
---

# Principles of Reproducibility

The idea:  You should be able to send your project to someone else, and they can jump in and start working right away.

1. Files are organized and well-named.
2. References to data and code work for everyone. (Projects!)
3. Package dependency is clear.  (`library`)
4. Code will run the same every time, even if data values change.
5. Analysis process is well-explained and easy to read.

---
# Bad practice

If you put something like this at the top of your .Rmd file, I will set your computer on fire:

```{r setwd, eval = FALSE}
setwd("/User/hglanz/Dropbox/Teaching/331/Data/")
```

- Setting working directory by hand = BAD!  
- That directory is specific to you!  Where should someone else look for the data?
- Markdown files won't knit this code.

---
class: center, middle
# Dealing with packages

---

## install.packages()

To use a package, it must first be *installed* to your copy of RStudio.

This only needs to happen **ONCE** per computer or server.

```{r install packages, eval = FALSE}
install.packages("tidyverse")
```

---

# update.packages()

If a computer you are using has the package, but you aren't sure if it's up to date:

```{r update packages, eval = FALSE}
update.packages("tidyverse")
```

to install the most recent version.

# NEVER EVER EVER EVER PUT "install.packages" or "update.packages" INTO YOUR MARKDOWN FILE!!!!!

---

# library() and require()

- A package must only be *installed* once.
- But it must be *loaded* every time, to tell R it is needed for this particular analysis.
- `library` gives an **error** if the package isn't installed
- `require` gives a `TRUE/FALSE` value depending on if the package was available
- For now, you should always use `library`

```{r library, eval = FALSE}
library(tidyverse)

# notice:  no quotes around package name
```


---
class: center, middle

# Deliverables in Markdown/Quarto/Notebooks

---
# General principles

- Give your file a good title, author, and date
- Name your file something meaningful (match the title!)
- Use `#` to make section headings etc.  (see cheatsheet)
- Each code chunk is a unified thought.
- Text before and after each code chunk to motivate your code and explain the output.
- Make deliberate decisions about what to display:
    + Does the reader need to see data loading step?
    + Does the reader need to see your code?
    + Does the reader need to see the output?
 
---   
# Naming code chunks

- Lets you jump between chunks - see demo.
- Careful with copy-pasting, duplicate names are not allowed.

---

# Setting chunk options

- `echo`: Should this code be **shown** (echoed)?
- `eval`: Should this code be **run** (evaluated)?
- `include`: Should this code be neither shown nor run?
- `error`: If there is an error, should that display rather than breaking the knit?
- `warning`: If there is a warning, should that display?
- `message`: If there is a message, should that display?

---

class: center, middle, invert

# YOUR TURN

---
class: center, middle

## Prettify your notebook with the penguin visualizations and add some narrative text/captions for your plots!